\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgements}

\renewcommand*\abstractname{Acknowledgements}

\begin{vplace}[0.7]

	\begin{abstract}

		\fontsize{11}{12}\selectfont\baselineskip=1.2em

		I would like to express my appreciation to Simone Gennai, Fabio Stella, Pietro Govoni and Glenn Dirkx, for giving me the chance to work on a project of this incredible magnitude, joining a unique community of passionate researchers. Furthermore, they gave me the freedom, the tools and the motivations to investigate my ideas, while also being critical and providing invaluable feedback.

		I extend my gratitude to everyone at CERN who gave me useful advices and precious insights: Dinyar Rabady, Andrea Bocci, Fabio Espinosa, Adrian Alan Pol, Alessandro Thea and Sam James Harper.

		Thanks to CERN and the CMS collaboration for providing the funds, the datasets and the computing resources used in this work.

		At my home university, I would like to mention Leonardo Mariani, Daniela Micucci, Edoardo Datteri and Roberto Previtera. I also thank the students I supervised, Riccardo, Alessandro, Morris, Luca and Stefano, from whom I learned a lot.

		I send my love to the inspiring and brilliant teachers I had during my earlier education: Camilla D'Andria, Caterina Brasacchio, Maurizia Perego, Luca Mauri.

		Thanks to my friends in Italy and France, too many to name them all, for keeping me sane, loved and happy.

		Last, but definitely not least, my family has been essential and I can't express how grateful I am for the support, the unconditional love and the appreciation with which they always enabled my passions and promoted my choices.

	\end{abstract}

\end{vplace}

\thispagestyle{empty}

